Discovering causal factors with back-and-forth regression

### 2019-05-16 paper/paper.tex
modified asymptotic proof

\subsection{Asymptotic behaviour - linear case}
As discussed in the introduction, our problem can be formulated (up to a redifinition of the features of X) as $Y=f(MEX+N)$, with F an unknown function.
%
If F is linearm we can rewrite it as $Y = F(MEX + N)$, with X an (N, dx) matrix of possible causes, E a (dx,dx) binary diagonal matrix selecting active features of X.
%
M an unknown full rank (dz,dx) transformation of EX, N a (dz,N) homoscedastic noise matrix, F an unknown (possibly full rank) (dy,dz) transformation corresponding to the measuring apparatus, and Y a (dx,dy) matrix of measured effects.

Since N is centred, the first least square regression (X from Y), amounts to finding
\begin{equation}
\begin{aligned}
\min_G \left \| X-GY \right \|^2 &= \min_G \left \| X - GF(MEX+N)\right\|^2 \\
&{}= \min_G \left \| (I-GFME)X\right\| ^2 + \left \| GFN\right \| ^2\\
&{}\leq \min_G \left \| I-GFME\right\| ^2 \left \| X\right\| ^2 + k\left \| G\right \| ^2\\
\end{aligned}
\end{equation}
with k a positive constant which depends on the variance of FN.
%
This amounts to finding the minimal norm generalized inverse of FME, or its Moore-Penrose inverse $\hat  G= (FME)^{\dagger}$.

The second regression is from X to the predicted value of X in the above, or $\hat G Y$
\begin{equation}
\begin{aligned}
arg \min_H \left \| (FME)^{\dagger}Y - HX \right \|^2 &= arg \min_H \left \| (FME)^{\dagger}F(MEX+N) - HX \right \|^2 \\
&= arg \min_H \left \| ((FME)^{\dagger}(FME)-H)X \right \| ^2 + \left \| (FME)^{\dagger}FN \right \| ^2\\
&= arg \min_H \left \| ((FME)^{\dagger}(FME)-H)X \right \| ^2\\
&=(FME)^{\dagger}(FME)
\end{aligned}
\end{equation}
Thus, in the linear case, double regression retrieves $(FME)^{\dagger}(FME)$.

For any matrix A, $A^\dagger A$ is the orthogonal projector over $Img(A')$, and $I-A^\dagger A$ the orthogonal projector over $Ker(A)$.
%
In our case, $Ker(FME)\supseteq Ker(E)$.
%
If FM has full rank over the subspace spanned by the features selected by E (this is assumed for M and will happen for almost any random F), then $Ker(FME) = Ker(E)$,.
%
This means $I - \hat H$ is the projector over $Ker(E)$, the diagonal matrix that spans the unselected features, and $\hat H$ is the diagonal matrix corresponding to the selected features, that is E.
%
Under such hypotheses, $\hat H$ retrieves E, the features of X that have a causal influence on Y.
%
Since it is diagonal with eigenvalues well separated (0 and 1), it should do so in a fairly resilient way.

\subsection{Asymptotic behaviour - non linear case}

>>>>

\subsection{Asymptotic behaviour - linear case}
As discussed in the introduction, our problem can be formulated (up to a redifinition of the features of X) as $Y=f(EX+N)$, with F an unknown function.
%
If F is linear we can rewrite it as $Y = F(EX + N)$, with X an (n, dx) matrix of possible causes, E a (dx,dx) binary diagonal matrix selecting active features of X.
%
N a (dx,n) homoscedastic noise matrix, F an unknown (possibly full rank) (dy,dx) transformation corresponding to the measuring apparatus, and Y a (n,dy) matrix of measured effects.

Since N is centred, the first least square regression (X from Y), amounts to finding
\begin{equation}
\begin{aligned}
\min_G \left \| X-GY \right \|^2 &= \min_G \left \| X - GF(EX+N)\right\|^2 \\
&{}= \min_G \left \| (I-GFE)X\right\| ^2 + \left \| GFN\right \| ^2\\
&{}= \min_G \alpha \left \| I-GFE\right\| ^2  + \beta \left \| G\right \| ^2\\
\end{aligned}
\end{equation}
with $\alpha$ and $\beta$ positive constants, the latter equal to the norm of FN. We are balancing two minimisation, with $/alpha$ close to zero, the solution is the null matrix, with $\beta$ close to zero, it is the pseudo inverse of $FE$.  
%

In the general case, $\beta=\left \| FN\right \| ^2$, and the solution is $k (FE)^\dagger$, with $k= {\alpha\over{\alpha+\beta}}$.

The second regression is from X to the predicted value of X in the above, or $\hat G Y$
\begin{equation}
\begin{aligned}
arg \min_H \left \| k(FE)^{\dagger}Y - HX \right \|^2 &= arg \min_H \left \| k(FE)^{\dagger}F(EX+N) - HX \right \|^2 \\
&= arg \min_H \left \| (k(FE)^{\dagger}(FE)-H)X \right \| ^2 + \left \| k(FE)^{\dagger}FN \right \| ^2\\
&= arg \min_H \left \| (k(FE)^{\dagger}(FE)-H)X \right \| ^2\\
&=k(FE)^{\dagger}(FE)
\end{aligned}
\end{equation}
Thus, in the linear case, double regression retrieves $k(FE)^{\dagger}(FE)$. In many cases, if v s the variance of $FN$, k can be approximated as $k= {1\over{1+v}}$, it quantifies the amount of noise added to the causal process, and the difficulty of recovering the causal features.

For any matrix A, $A^\dagger A$ is the orthogonal projector over $Img(A')$, and $I-A^\dagger A$ the orthogonal projector over $Ker(A)$.
%
In our case, $Ker(FE)\supseteq Ker(E)$.
%
If F has full rank over $Img(E)$, the subspace spanned by E, then $Ker(FE) = Ker(E)$,.
%
This means $I - \hat H$ is the projector over $Ker(E)$, the diagonal matrix that spans the unselected features, and $\hat H$ is k times the diagonal matrix corresponding to the selected features, that is kE. Since it is diagonal with eigenvalues 0 and k, if k is large enough, it should allow us to retrieve causal features in a fairly robust way.
\subsection{Asymptotic behaviour - non linear case}
